---
title: "Life Expectancy & SDOH"
author: "Bertram Melix"
date: "6/03/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Library Packages Required for Analysis
```{r}
library("rgeos")
library("rgdal")
library("spdep")
library("maptools")
library("MASS")
library("robust")
library("faraway")
library("spatialreg")
library("car")
```


## Read shapefile with rgdal
```{r}
FlShp <- rgdal::readOGR(dsn = "C:/Users/Bertram/Desktop/LifeExpectancy/Shapefile", layer = "LESDOH")
```


## List column headers
```{r}
names(FlShp)
```


## Create queen neighbor file
```{r}
LEQueen<- poly2nb(FlShp,queen=TRUE)
```


## Convert queen neighbor to list
```{r}
LEQueen1<-nb2listw(LEQueen, zero.policy = TRUE)
```


## Test for Multicollinearity with VIF function
```{r}
TractsVIF <- rgdal::readOGR(dsn = "C:/Users/Bertram/Desktop/LifeExpectancy/Shapefile", layer = "LESDOHVIF")
names(TractsVIF) 
LEVIF <- data.frame(TractsVIF[, c(2,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34)])
names(LEVIF)
VIFModel<- lm(LE_Estimat~., LEVIF)
summary(VIFModel)
faraway::vif(VIFModel)
```


## Formula
```{r}
LEform1<- formula(LE_Estimat~QBLACK+QNATAM+QASIAN+QHISP+QAGEDEP+PPUNIT+QRENTER+QNRRES+QFEMALE+QFHH+PERCAP+QESL+QCVLUN+QPOVTY+QMOHO+QED12LES+QFEMLBR+QEXTRCT+QSERV+QNOAUTO+QFAM+QRICH200K+QSSBEN+QURBAN+POPDENS+QUNINS)
```


## OLS Model
```{r}
LEFitlm<- lm(LEform1, data= FlShp)
summary(LEFitlm)
LEResFitlm<- resid(LEFitlm)
stepAIC(LEFitlm, direction="backward")
```


## QQPlot
```{r}
car::qqPlot(resid(LEFitlm))
pred_quants <- qnorm((rank(resid(LEFitlm)) - 0.5) / length(resid(LEFitlm)))
points(resid(LEFitlm) ~ pred_quants, pch = 1, col = "red")
```


## Fitted values vs Residuals, QQPlot 
```{r}
plot(LEFitlm)
```


## Shapiro Wilkes Test
```{r}
shapiro.test(resid(LEFitlm))
```


## Lagrange Multiplier Test
```{r}
lm.LMtests(LEFitlm,LEQueen1,zero.policy = TRUE, test = c("LMerr","RLMerr","LMlag","RLMlag"))
```


## Moran's I test on residuals from Linear Regression
```{r}
moran.test(LEFitlm$residuals, listw = LEQueen1, zero.policy = TRUE)
```


## Spatial Durbin Error Model (SDEM)
```{r}
LEFitErrorM<- errorsarlm(LEform1, data=SE, listw=LEQueen1, etype = "emixed", tol.solve=1.0e-30, zero.policy = TRUE)
summary(LEFitErrorM)

```


## Moran's I on residuals from SDEM
```{r}
moran.test(LEFitErrorM$residuals, LEQueen1, zero.policy = TRUE)
```


## Spatial Breusch-Pagan test for Heterskadisticity
```{r}
bptest.sarlm(LEFitErrorM, studentize = TRUE)
```


## Psuedo R^2
```{r}
1-(LEFitErrorM$SSE)/(var(FlShp$LE_Estimat)*(length(FlShp$LE_Estimat)-1))
```


## AIC
```{r}
AIC(LEFitlm,LEFitErrorM)
```

